---
title: "Predict MPG based on HP, time to 1:4 mile, and transmission type"
runtime: shiny
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Model design

* An analysis of mtcars dataset has been conducted in order to design a regression model that could successfully explain variations in vehicle mileage 

* The analysis revealed that optimal regression parameters predicting gas mileage include:
    + weight (wt)
    + time to 1/4 mile (qsec) 
    + type of transmission (auto or manual)

* A model has been fitted to predict the outcome based on these three parameters. Final model is represented by mpg ~ wt + qsec + am


## Instructions
- The application takes the input for the three model parameters (weight, time to 1/4 mile and type of transmission)

- The application predicts the gas mileage for a vehicle with the parameters passed as an input. 

- In addition, the application builds plots for the two continuous predictors against mileage (mpg against wt, and mpg against qsec) while showing factor transmission predictor as color.

- All the vehicles in the dataset are plotted to indicate the relative position of the input parameters to those of real vehicles.


## Predict MPG
```{r}
shinyApp(
  
  ui = fluidPage(
  
  # Application title
  
  
  # Sidebar with a slider input for number of bins 
  sidebarLayout(
    sidebarPanel(
      sliderInput("SliderWeight",
                  "What is the weight of this car (in 1000 lbs)?",
                  min = 1.5,
                  max = 5.5,
                  value = 3.5),
      sliderInput("SliderQsec",
                  "How fast does this car does 1/4 mile (in sec)?",
                  min = 13,
                  max = 23,
                  value = 18),
      checkboxInput("auto_manual", "Check if transmission is automatic", value = TRUE)
      
     
     
    ),
    
    # Show a plot of the generated distribution
    mainPanel(
      h3("Predicted mpg:"), 
      textOutput("pred1"),
      plotOutput("Plot1", height = "240px"),
      plotOutput("Plot2", height = "240px")
      
      
      
    )
  )
),
  
  server = function(input, output) {
  
  model1 <- lm(mpg ~ wt + qsec + am, data = mtcars)
  cars <- mtcars
  cars$am <- as.factor(cars$am)
  
  model1pred <- reactive({
    wtInput <- input$SliderWeight
    qsecInput <- input$SliderQsec
    amInput <- ifelse(input$auto_manual, 0, 1)
    predict(model1, newdata = data.frame(wt = wtInput, qsec = qsecInput, am = amInput))
  })
  
  
  output$Plot1 <- renderPlot({
    wtInput <- input$SliderWeight
    qsecInput <- input$SliderQsec
    amInput <- ifelse(input$auto_manual, 1, 0)
    plot(mtcars$wt, mtcars$mpg, main = "MPG dependency on weight", xlab = "weight", ylab = "mpg", pch = 16, xlim = c(1.5, 5.5),
         ylim = c(10, 35), col = cars$am)
    points(wtInput, model1pred(), col = ifelse(amInput == 1, "red", "black"), pch = 16, cex = 4)
    legend("topright", legend = c("manual", "auto"), fill = c("black", "red"), box.lty = 0)
  })
  
  output$Plot2 <- renderPlot({
    wtInput <- input$SliderWeight
    qsecInput <- input$SliderQsec
    amInput <- ifelse(input$auto_manual, 1, 0)
    plot(mtcars$qsec, mtcars$mpg, main = "MPG dependency on time to 1/4 mile", xlab = "qsec", ylab = "mpg", pch = 16, xlim = c(13, 23),
         ylim = c(10, 35), col = cars$am)
    points(qsecInput, model1pred(), col = ifelse(amInput == 1, "red", "black"), pch = 16, cex = 4)
    legend("topright", legend = c("manual", "auto"), fill = c("black", "red"), box.lty = 0)
  })
  
  output$pred1 <- renderText({model1pred()})
  
},
  
  options = list(height = 500)
)

```


## R Output

- The model makes it possible to predict car model mileage with reasonable accuracy
- Assuming that car weight is typically available, it only needs one simple meausurement that doesn't require any special tools
- It doesn't require any specialized knowledge to be applied and the inference can be performed by anybody and anywhere with an access to browser

